### 2018-01-27 manuscript.rmd
incorporating final comments

---
bibliography: manuscript.bib
csl: molecular-phylogenetics-and-evolution.csl
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    includes:
      in_header: mystyles.sty
---

@@

All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017). The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag XXXXXXXXXX.

@@

Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig. 3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores including *Rudjakovia* sp. and (iii) monophyletic monoecious with dioecious *Rudjakovia* sp. <!--  XXX This needs to be fixed to coincide with the two hypotheses listed in section 2.3 XXX -->. In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig. S2).

@@

Bracts are greatly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica* (Fig. 1K), and in *Gymnopraia lapislazula*. These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny. In species without bracts, other zooids appear to fulfill the roles of neutral buoyancy and protection. In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis]. *Hippopodius hippopus* has up to twelve nectophores and can retract the stem within the nectophores. The nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig. 1C shows the blanching of nectophores)[@bassot1978bioluminescence].  <!-- XXX This hippo section seems like a total non-sequitur in the bract paragraph, even considering the potantial role in buoyancy XXX -->

@@

![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)

@@

![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](sowh/sowh_tests_supplemental_figure/sowh_tests_supplemental_figure.png)

@@

```{r tentilla, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla."}
#Simmap TENTILLA
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
plot(tentilla_obj)
```

```{r tentilla2, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla."}
#Simmap TENTILLA
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
plot(tentilla2_obj)
```

```{r nectosome, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
#Simmap NECTOSOME
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
plot(nectosome_obj)
```

```{r desc_mantle, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores. Cystonects and Athorybia were excluded as they do not have a nectosome. "}
#Simmap Descending Mantle Canal
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
plot(desc_mantle_obj)
```

```{r nectosome_pos, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome. Cystonects were excluded as they do not have a nectosome."}
#SIMMAP NectoPos
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Dorsal", "Ventral")
# necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
# nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
plot(necto_pos_obj)
```

```{r median_cont, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
depth_value = depth_median
depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
plot(depthmap)
```

@@

>>>>

---
bibliography: manuscript.bib
csl: molecular-phylogenetics-and-evolution.csl
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: mystyles.sty
  word_document: default
---

@@

All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017). The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v1".

@@

Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following two alternative phylogenetic hypotheses against the most likely tree topology (Fig. 3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores. As the sexual mode of  *Rudjakovia* is undescribed, we carried out two tests: with one test including *Rudjakovia* sp. as a monoecious species, and the other with dioecious *Rudjakovia* sp. In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig. S2).

@@

Bracts are greatly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica* (Fig. 1K), and in *Gymnopraia lapislazula*. These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny. In species without bracts, other zooids appear to fulfill the roles of neutral buoyancy and protection. In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis]. While in *Hippopodius hippopus* the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig. 1C shows the blanching of nectophores)[@bassot1978bioluminescence].

@@

![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)

@@

![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](supplementary_info/Supplementary_Figure2.png)

@@

```{r tentilla, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla."}
#Simmap TENTILLA
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
plot(tentilla_obj)

pdf("supplementary_info/Supplementary_Figure3.pdf")
plot(tentilla_obj)
dev.off()

```{r tentilla2, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla."}
#Simmap TENTILLA
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
plot(tentilla2_obj)

pdf("supplementary_info/Supplementary_Figure4.pdf")
plot(tentilla2_obj)
dev.off()
```

```{r nectosome, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
#Simmap NECTOSOME
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
plot(nectosome_obj)

pdf("supplementary_info/Supplementary_Figure5.pdf")
plot(nectosome_obj)
dev.off()
```

```{r desc_mantle, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores. Cystonects and Athorybia were excluded as they do not have a nectosome. "}
#Simmap Descending Mantle Canal
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Absent", "Present")
# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
plot(desc_mantle_obj)

pdf("supplementary_info/Supplementary_Figure6.pdf")
plot(desc_mantle_obj)
dev.off()
```

```{r nectosome_pos, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome. Cystonects were excluded as they do not have a nectosome."}
#SIMMAP NectoPos
# plotTree(ultratree, lwd = 4)
# colors = c("black", "red")
# names(colors) = c("Dorsal", "Ventral")
# necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
# nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
plot(necto_pos_obj)

pdf("supplementary_info/Supplementary_Figure7.pdf")
plot(necto_pos_obj)
dev.off()
```

```{r median_cont, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
depth_value = depth_median
depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
plot(depthmap)

pdf("supplementary_info/Supplementary_Figure8.pdf")
plot(depthmap)
dev.off()
```

@@

pdf("supplementary_info/Supplementary_Figure9.pdf")
plot(depthmap_ext)
dev.off()
